---
title: "Predicting Bank Customer Churn using AWS SageMaker and XGBoost in Local RStudio"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

theme_dark_grey <- function(base_size=14, base_family="sans") {
   library(grid)
   library(ggthemes)
   (theme_foundation(base_size=base_size, base_family=base_family)
      + theme(plot.title = element_text(face = "bold", colour = '#ffffb3',
                                        size = rel(1.1), hjust = 0.0, 
                                        margin = margin(0,0,5,0)),
              text = element_text(),
              panel.background = element_rect(colour = NA, fill = 'grey20'),
              plot.background = element_rect(colour = NA, fill = '#262626'),
              panel.border = element_rect(colour = NA),
              axis.title = element_text(face = "bold",size = rel(1), colour = 'white'),
              axis.title.y = element_text(angle=90,vjust =2),
              axis.title.x = element_text(vjust = -0.2),
              axis.text = element_text(colour = 'grey70'), 
              axis.line.x = element_line(colour="grey70"),
              axis.line.y = element_line(colour="grey70"),
              axis.ticks = element_line(colour="grey70"),
              panel.grid.major = element_line(colour="#262626"),
              panel.grid.minor = element_blank(),
              legend.background = element_rect(fill ='#262626'),
              legend.text = element_text(color = 'white'),
              legend.key = element_rect(colour = NA, fill = '#262626'),
              legend.position = "bottom",
              legend.direction = "horizontal",
              legend.box = "vetical",
              legend.key.size= unit(0.5, "cm"),
              #legend.margin = unit(0, "cm"),
              legend.title = element_text(face="italic", colour = 'white'),
              plot.margin=unit(c(5,5,5,5),"mm"),
              strip.background=element_rect(colour="#2D3A4C",fill="#2D3A4C"),
              strip.text = element_text(face="bold", colour = 'white')
      ))
}

scale_fill_todd_dark <- function(...){
   library(scales)
   discrete_scale("fill","todd",manual_pal(values = c("#decbe4","#fed9a6","#b3cde3","#ccebc5","#fbb4ae",
                                                      "#ffffcc","#e5d8bd","#fddaec","#f2f2f2")), ...)
}
```

```{r echo=TRUE}
library(tidymodels)
library(tidyverse)
library(scales)
library(lubridate)
library(stacks)
library(corrmorant)
library(patchwork)
library(reticulate)    # for calling the SageMaker Python SDK from R
library(pROC)          
```

-   [ðŸ¥… Project Goal](#goal)
-   [ðŸ—‚ Obtain Data](#data)
-   [ðŸ› Clean Data](#clean)
-   [ðŸ”­ Explore Data](#explore)
-   [ðŸ§® Prep XGBoost](#prep)
-   [ðŸ— Build Workflows](#build)
-   [ðŸ”§ Tune Models](#tune)
-   [ðŸ† Forecast w/ Best Models](#forecast)
-   [ðŸ Ensemble & Save Work](#ensemble)
-   [ðŸ¤” Decisions](#decisions)

## ðŸ¥… Goal of this Project {#goal}

Predict whether a bank customer will churn using AWS SageMaker and RStudio

## ðŸ—‚ Obtain Data {#data}

The data comes from the Kaggle competition
[SLICED](https://www.kaggle.com/c/sliced-s01e07-HmPsw2/overview), a live data science
competition on Twitch where competitors get a never-before-seen dataset and two-hours
to code a solution to a prediction challenge.

```{r message=FALSE, warning=FALSE}
churn_data <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-sagemaker-r-xgb-churn/train.csv")


holdout <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-sagemaker-r-xgb-churn/test.csv")

```

## ðŸ› Clean Data {#clean}

```{r }
glimpse(churn_data)
```

### [Data Dictionary]{.ul}

#### Outcome Variable

-   attrition_flag: whether the customer is churned (0 = no; 1 = yes)

#### ID

-   id: unique identifier for the customer

#### Categorical Features

-   gender
-   education_level
-   income_category: income range of the customer

#### Numeric Features

-   customer_age

-   total_relationship_count: number of relationships

-   months_inactive_12_mon: number of months the customer is inactive in past 12
    months

-   credit_limit

-   total_revolving_bal: customer's total revolving balance

-   total_amt_chng_q4_q1: amount the balance changed from Q4 to Q1

-   total_trans_amt: value of all the customer's transactions in the period

-   total_trans_ct: count of all the customer's transactions

-   total_ct_chng_q4_q1: difference in number of the customer's transactions from Q4
    to Q1

-   avg_utilization_ratio: customer's average utilization ratio during the period

```{r message=FALSE, warning=FALSE, include=FALSE}
DataExplorer::create_report(churn_data)
```

```{r}
skimr::skim_without_charts(churn_data)
```

```{r message=FALSE, warning=FALSE}
churn_data2 <- churn_data %>% 
  mutate(churned = factor(ifelse(attrition_flag == 1, "yes", "no"),
                          levels = c("no", "yes"))) %>%
  mutate(education_level = fct_relevel(education_level, c("Unknown", 
                                                          "Uneducated", 
                                                          "High School", 
                                                          "College", 
                                                          "Graduate", 
                                                          "Doctorate", 
                                                          "Post-Graduate"))) %>%
  mutate(income_category = fct_relevel(income_category, c("Unknown",
                                                          "Less than $40K",
                                                          "$40K - $60K",
                                                          "$60K - $80K",
                                                          "$80K - $120K",
                                                          "$120K +")))

churn_data3 <- churn_data2 %>% 
               mutate(across(.cols = c(credit_limit, total_trans_amt),
                             .fns  = log1p)) %>%
               mutate(across(.cols = c(credit_limit, total_trans_amt), 
                             .fns = timetk::standardize_vec)) %>% 
               select(-c(id, attrition_flag))
```

```{r message=FALSE, warning=FALSE}

corfun <- function(x, y)  {
  round(cor(x, y, method = "pearson", use = "pairwise.complete.obs"), 2)
}

ggcorrm(churn_data3, aes(col = churned, fill = churned), bg_dia  = "grey30") +
    lotri(geom_point(alpha = 0.1)) +
    lotri(geom_smooth(se=F, method = "loess")) +
    utri_funtext(fun = corfun, size = 6) +
    dia_names(y_pos = 0.15, size = 3) +
    dia_density(lower = 0.3, fill = "grey60", color = 1) +
    theme_dark() +
    scale_color_viridis_d() +
    scale_fill_viridis_d() +
    labs(title = "Correlation Plot")
```

```{r}
dense <- churn_data3 %>%
         select(-c(gender, education_level, income_category)) %>%
         pivot_longer(cols = c(customer_age:avg_utilization_ratio), 
                      names_to = "feature", 
                      values_to = "value")

ggplot(dense, aes(value, fill = churned)) +
  geom_density(alpha = .5) +
  facet_wrap(~ feature, scales = "free") +
  labs(title = "Numeric features impacting churn?")
```

```{r}
churn_data %>% 
  ggplot(aes(total_trans_amt, total_trans_ct, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()

churn_data %>% 
  ggplot(aes(total_ct_chng_q4_q1, total_trans_ct, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()

churn_data %>% 
  ggplot(aes(credit_limit, avg_utilization_ratio, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()

churn_data %>% 
  ggplot(aes(total_trans_amt, total_amt_chng_q4_q1, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 60) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()
```

```{r echo=FALSE}
cat_feat <- churn_data2 %>% 
  select(c(gender, education_level, income_category, churned))

n_gen <- cat_feat %>% 
  group_by(gender) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(gender, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none',
              axis.title.x = element_blank())

n_edu <- cat_feat %>% 
  group_by(education_level) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(education_level, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

n_inc <- cat_feat %>% 
  group_by(income_category) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(income_category, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

n_chrn <- cat_feat %>% 
  group_by(churned) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(churned, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none',
              axis.title.x = element_blank())

n_chrn + n_gen + n_edu + n_inc +
plot_layout(ncol = 2, heights = c(0.7,2)) +
plot_annotation(title = "Counts of Each Categorical Variable in Training Data")
```

```{r echo=FALSE}
gen <- ggplot(cat_feat %>% 
                group_by(gender, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
               aes(gender, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                  color    = "white",
                  size     = 4,
                  fontface = 'bold',
                  hjust    = 1.3,
                  vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

ed <- ggplot(cat_feat %>% 
                group_by(education_level, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
             aes(education_level, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                    color    = "white",
                    size     = 4,
                    fontface = 'bold',
                    hjust    = 1.3,
                    vjust    = 0.4) +
        theme_minimal() +
        theme(axis.title.x = element_blank(),
              legend.position = 'none')

inc <- ggplot(cat_feat %>% 
                group_by(income_category, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
              aes(income_category, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                    color    = "white",
                    size     = 4,
                    fontface = 'bold',
                    hjust    = 1.3,
                    vjust    = 0.4) +
        theme_minimal() +
        theme(axis.title.x = element_blank(),
              legend.position = 'none')

ed + inc + gen +
    plot_layout(ncol = 2, heights = c(2,0.7)) +
    plot_annotation(title = "Percentage of Each Categorical Variable that Churned")

```

```{r echo=FALSE}
cat_feat %>%
  pivot_longer(gender:income_category) %>%
  ggplot(aes(y = value, fill = churned)) +
  geom_bar(position = "fill") +
  facet_wrap(vars(name), scales = "free", ncol = 1) +
  theme_minimal() +
  labs(x = NULL, y = NULL)
```

final_fit %\>% vip(geom = "col", aesthetics = list(fill = "midnightblue", alpha =
0.8)) + scale_y\_continuous(expand = c(0, 0))

or

workflow() %\>% add_recipe(ranger_recipe) %\>% add_model(imp_spec) %\>%
fit(ikea_train) %\>% pull_workflow_fit() %\>% vip(aesthetics = list(alpha = 0.8, fill
= "midnightblue"))

```{r}
use_condaenv("sagemaker-r", required = TRUE)
sagemaker <- import("sagemaker")
session <- sagemaker$Session()
```

```{r}
# bucket <- session$default_bucket()
bucket <- "twarczak-sagemaker2"
project <- "bank-churn"
data_path <- paste0("s3://", bucket, "/", project, "/", "data")
models_path <- paste0("s3://", bucket, "/", project, "/", "models")
```

```{r}
set.seed(333)
splits  <- initial_split(churn_data2, prop = 0.85, strata = churned)
churn_training <- training(splits)
churn_validation  <- testing(splits)
churn_test <- holdout
```

```{r}
churn_rec <- recipe(attrition_flag ~ ., data = churn_training) %>% 
             step_rm(churned) %>% 
             update_role(id, new_role = "id variable") %>%
             step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE)

churn_preprocessed <- prep(churn_rec, training = churn_training) %>% 
                      juice() %>% 
                      select(attrition_flag, everything()) # just to put attrition_flag as 1st column

dim(churn_preprocessed)
```
```{r}

```


```{r}
s3_uploader <- sagemaker$s3$S3Uploader()

s3_train <- s3_uploader$upload(local_path = "../bank-churn/data/churn_training.csv", 
                               desired_s3_uri = data_path)

s3_validation <- s3_uploader$upload(local_path = "../bank-churn/data/hotels_validation.csv",
                                    desired_s3_uri = data_path)

s3_test <- s3_uploader$upload(local_path = "../xgboost-hotels/data/hotels_test.csv", 
                              desired_s3_uri = data_path)
```

```{r}
region <- session$boto_region_name
# get container image location
container <- sagemaker$image_uris$retrieve(framework = "xgboost", 
                                           region    = region, 
                                           version   = "1.3-1" )
# get SageMaker execution role stored in .Renviron
role_arn <- Sys.getenv("SAGEMAKER_ROLE_ARN")
#role_name <- "AmazonSageMaker-ExecutionRole-20210727T164660"
# Create an Estimator object
xgb_estimator <- sagemaker$estimator$Estimator(image_uri          = container,
                                               role               = role_arn,
                                               instance_count     = 1L,
                                               instance_type      = "ml.m5.4xlarge",
                                               volume_size        = 30L,
                                               max_run            = 3600L,
                                               output_path        = models_path,
                                               sagemaker_session  = session,
                                               use_spot_instances = TRUE,
                                               max_wait           = 600L )
```

```{r}
xgb_estimator$set_hyperparameters(objective        = "binary:logistic",
                                  eval_metric      = "auc",
                                  max_depth        = 5L, 
                                  eta              = 0.1,
                                  num_round        = 100L,
                                  colsample_bytree = 0.4,
                                  alpha            = 10L,
                                  min_child_weight = 1.1,
                                  subsample        = 0.7)
```

```{r}
# Create training job name based project organization principles
algo <- "xgboost"
timestamp <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
job_name <- paste(project, algo, timestamp, sep = "-")
s3_train_input <- sagemaker$inputs$TrainingInput(s3_data = s3_train,
                                                 content_type = 'csv')
s3_valid_input <- sagemaker$inputs$TrainingInput(s3_data = s3_validation,
                                                 content_type = 'csv')
input_data <- list('train'      = s3_train_input,
                   'validation' = s3_valid_input)
```

```{r}
xgb_estimator$fit(inputs = input_data,
                  job_name = job_name,
                  wait = FALSE)  # If set to TRUE, the call will wait until job completes
```

```{r}
session$describe_training_job(job_name)[["TrainingJobStatus"]]
```

```{r}
training_job_stats <- session$describe_training_job(job_name = job_name)
final_metrics <-  map_df(training_job_stats$FinalMetricDataList, 
                          ~tibble(metric_name = .x[["MetricName"]],
                                  value = .x[["Value"]]))
final_metrics
```

```{r}
predictions_path <- paste0(models_path, "/", job_name, "/predictions")
xgb_batch_predictor <- xgb_estimator$transformer(instance_count = 1L, 
                                                 instance_type  = "ml.m5.xlarge", 
                                                 strategy       = "MultiRecord",
                                                 assemble_with  = "Line",
                                                 output_path    = predictions_path)
```

```{r}
xgb_batch_predictor$transform(data         = s3_test, 
                              content_type = 'text/csv',
                              split_type   = "Line",
                              job_name     = job_name,
                              wait         = FALSE) # If TRUE, call waits until job completes
```

```{r}
session$describe_transform_job(job_name)[["TransformJobStatus"]]
```

```{r}
s3_downloader <- sagemaker$s3$S3Downloader()
s3_test_predictions_path <- s3_downloader$list(predictions_path)
 
dir.create("./predictions")
s3_downloader$download( s3_test_predictions_path, "./predictions")
 
test_predictions <- read_csv("./predictions/hotels_test.csv.out",
                              col_names = FALSE) %>% 
                    pull(X1)

test_results <- tibble(truth       = hotels_test$children,
                       predictions = test_predictions )

head(test_results)
```

```{r}
roc_obj <- roc(test_results$truth,
               test_results$predictions,
               plot        = TRUE,         
               grid        = TRUE,
               print.auc   = TRUE,
               legacy.axes = TRUE, 
               main        = "ROC curve for XGBoost classification",
               show.thres  = TRUE,
               col         = "red2" )
```

```{r}
conf_matrix <- confusionMatrix(factor(ifelse(test_results$predictions >= 0.5, 1, 0),
                                      levels = c("0", "1"), 
                                      labels = c("no children", "children")),
                               factor(test_results$truth, 
                                      levels = c(0, 1), 
                                      labels = c("no children", "children")),
                               positive = "children")

conf_matrix
```

```{r}
test_results_pred <- test_results %>% 
                     mutate(pred = factor(ifelse(predictions >= 0.5, 1, 0), 
                                          levels = c("0", "1")))

gmodels::CrossTable(x          = test_results_pred$pred, 
                    y          = test_results_pred$truth,
                    prop.chisq = FALSE, 
                    prop.r     = FALSE, 
                    prop.c     = FALSE, 
                    prop.t     = FALSE,
                    dnn        = c('predicted children', 'actual children'))
```

```{r}
```

```{r}
```

```{r}
```
