---
title: "Predicting Bank Customer Churn using AWS SageMaker and XGBoost in Local RStudio"
output: md_document
author: Todd Warczak
date: '2021-08-03'
slug: modeltime
categories:
   - rstats
   - AWS
   - SageMaker
   - XGBoost
   - R
tags:
   - rstats
   - AWS
   - reticulate
   - XGBoost
   - R
summary: "For this post, I experimented using AWS SageMaker with the AWS built-in XGBoost algorithm from within my local RStudio to predict whether a bank customer has churned. The data comes from the SLICED season 1 episode 7 Kaggle competition. SLICED is a data science competition where contestants are given a never-before-seen dataset and two-hours to code a solution to a prection challenge."
featured: "featured-sagemaker.png"
image:
  caption: ''
  focal_point: ''
  preview_only: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo=TRUE}
library(tidymodels)
library(tidyverse)
library(corrmorant)    # correlation matrix
library(patchwork)     # multiple plots to 1 plot
library(reticulate)    # calling the SageMaker Python SDK from R
library(pROC)          # ROC curves
library(viridis)       # color palletes
library(caret)         # confusion matrix
```

-   [🥅 Project Goal](#goal)
-   [🗂 Obtain Data](#data)
-   [🛁 Clean Data](#clean)
-   [🔭 Explore Data](#explore)
-   [🧮 AWS Setup](#aws)
-   [🏗 Train XGBoost](#build)
-   [🔧 Tune XGBoost](#tune)
-   [🏆 Select Best Model](#best)
-   [🎯 Predict Holdout](#holdout)
-   [📬 Submission](#submit)

## 🥅 Goal of this Project {#goal}

Predict whether a bank customer will churn using AWS SageMaker and RStudio

## 🗂 Obtain Data {#data}

The data comes from the Kaggle competition
[SLICED](https://www.kaggle.com/c/sliced-s01e07-HmPsw2/overview).

```{r chunk3, message=FALSE, warning=FALSE}
churn_data <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-sagemaker-r-xgb-churn/train.csv")

holdout <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-sagemaker-r-xgb-churn/test.csv")
```

`holdout` does not contain `attrition_flag`. Use `holdout` to run inference after
you've selected the best tuned model and built an endpoint to host the model. To
submit predicted customer churn to the SLICED competition, we'll need a .csv file
that contains only 2 columns: "id" (those found in `holdout`) and inferred
`attrition_flag`. That will be our final step.

## 🛁 Clean Data {#clean}

```{r chunk4, cache=TRUE}
glimpse(churn_data)
```

Let's look at the data dictionary to check we understand what each variable
represents.

### [Data Dictionary]

#### Outcome Variable

-   attrition_flag: whether the customer is churned (0 = no; 1 = yes)

#### ID

-   id: unique identifier for the customer

#### Categorical Features

-   gender
-   education_level
-   income_category: income range of the customer

#### Numeric Features

-   customer_age

-   total_relationship_count: number of relationships

-   months_inactive_12_mon: number of months the customer is inactive in past 12
    months

-   credit_limit

-   total_revolving_bal: customer's total revolving balance

-   total_amt_chng_q4_q1: amount the balance changed from Q4 to Q1

-   total_trans_amt: value of all the customer's transactions in the period

-   total_trans_ct: count of all the customer's transactions

-   total_ct_chng_q4_q1: difference in number of the customer's transactions from Q4
    to Q1

-   avg_utilization_ratio: customer's average utilization ratio during the period

I like using the `{skimr}` package to quickly check for missing values and get
variable information.

```{r chunk5, cache=TRUE}
skimr::skim_without_charts(churn_data)
```

Not much cleaning to do. No missing data. For EDA I want the `attrition_flag`
variable to be obvious whether the customer has churned or not. I'll factor the
categorical variables and experiment with log transforming some numeric variables.

## 🔭 Explore Data {#explore}

```{r chunk6, message=FALSE, warning=FALSE}
churn_data2 <- churn_data %>% 
  mutate(churned = factor(ifelse(attrition_flag == 1, "yes", "no"),
                          levels = c("no", "yes"))) %>%
  mutate(education_level = fct_relevel(education_level, c("Unknown", 
                                                          "Uneducated", 
                                                          "High School", 
                                                          "College", 
                                                          "Graduate", 
                                                          "Doctorate", 
                                                          "Post-Graduate"))) %>%
  mutate(income_category = fct_relevel(income_category, c("Unknown",
                                                          "Less than $40K",
                                                          "$40K - $60K",
                                                          "$60K - $80K",
                                                          "$80K - $120K",
                                                          "$120K +")))

churn_data3 <- churn_data2 %>% 
               mutate(across(.cols = c(credit_limit, total_trans_amt),
                             .fns  = log1p)) %>%
               mutate(across(.cols = c(credit_limit, total_trans_amt), 
                             .fns = timetk::standardize_vec)) %>% 
               select(-c(id, attrition_flag))
```

Let's look at a correlation matrix of all the numeric variables. We can group by our
new `churned` variable to see features that correlate differently based on the churn
outcome.

```{r chunk7, fig.width=12, fig.height=10, message=FALSE, warning=FALSE, cache=TRUE}

corfun <- function(x, y)  {
  round(cor(x, y, method = "pearson", use = "pairwise.complete.obs"), 2)
}

ggcorrm(churn_data3, aes(col = churned, fill = churned), bg_dia  = "grey30") +
    lotri(geom_point(alpha = 0.1)) +
    lotri(geom_smooth(se=F, method = "loess")) +
    utri_funtext(fun = corfun, size = 4) +
    dia_names(y_pos = 0.15, size = 2) +
    dia_density(lower = 0.3, fill = "grey60", color = 1) +
    theme_dark() +
    scale_color_viridis_d() +
    scale_fill_viridis_d() +
    labs(title = "Correlation Plot")
```

It looks to me that `total_ct_chng_q4_q1`, `total_trans_amt`, `total_trans_ct`,
`credit_limit`, and `avg_utilization_ratio` might be important to churn. I want to
see how the numeric feature density plots look and see if we can see any
distributions that are clearly separated by churn.

```{r chunk8, cache=TRUE}
dense <- churn_data3 %>%
         select(-c(gender, education_level, income_category)) %>%
         pivot_longer(cols = c(customer_age:avg_utilization_ratio), 
                      names_to = "feature", 
                      values_to = "value")

ggplot(dense, aes(value, fill = churned)) +
  geom_density(alpha = .5) +
  facet_wrap(~ feature, scales = "free") +
  labs(title = "Numeric features impacting churn?")
```

We see some clues in the distributions of `total_trans_amt`, `total_trans_ct`,
`total_relationship_count`, and `total_revolving_bal`. `months_inactive_12_mon` might
be important, but mostly when the variable = 1 month.

Let's try and find some combinations of these features that lead to churn more often.
Last data viz for the numeric features.

```{r chunk9.1, cache=TRUE}
churn_data %>% 
  ggplot(aes(total_trans_amt, total_trans_ct, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()
```

```{r chunk9.2, cache=TRUE}
churn_data %>% 
  ggplot(aes(total_ct_chng_q4_q1, total_trans_ct, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()
```

```{r chunk9.3, cache=TRUE}
churn_data %>% 
  ggplot(aes(credit_limit, avg_utilization_ratio, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 50) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark() 
```

```{r chunk9.4, cache=TRUE}
churn_data %>% 
  ggplot(aes(total_trans_amt, total_amt_chng_q4_q1, z = attrition_flag)) +
  stat_summary_hex(alpha = 0.8, bins = 60) +
  scale_fill_viridis_c() +
  labs(fill = "% Churned") +
  theme_dark()
```

It's clear that churn will be predicted if a customer data contains certain
combination of feature values (yellow).

Now let's get a count for each categorical variable and a percentage churn for each
category to see how balanced the dataset is.

```{r chunk10, cache=TRUE}
cat_feat <- churn_data2 %>% 
  select(c(gender, education_level, income_category, churned))

n_gen <- cat_feat %>% 
  group_by(gender) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(gender, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none',
              axis.title.x = element_blank())

n_edu <- cat_feat %>% 
  group_by(education_level) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(education_level, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

n_inc <- cat_feat %>% 
  group_by(income_category) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(income_category, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

n_chrn <- cat_feat %>% 
  group_by(churned) %>% 
  summarise(count = n()) %>% 
  ggplot(aes(churned, count)) +
  geom_col() +
  coord_flip() +
  geom_text(aes(label = count),
            color    = "white",
            size     = 3,
            fontface = 'bold',
            hjust    = 1.2,
            vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none',
              axis.title.x = element_blank())

n_chrn + n_gen + n_edu + n_inc +
plot_layout(ncol = 2, heights = c(0.7,2)) +
plot_annotation(title = "Counts of Each Categorical Variable in Training Data")
```

Our churn variable is unbalanced. Would be better if we had more data points for
those that have churned. We could upsample positive churn outcome or downsample the
negative churn outcome, but I'll let it slide this time.\

How balanced are the categorical predictors?

```{r chunk11, cache=TRUE}
gen <- ggplot(cat_feat %>% 
                group_by(gender, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
               aes(gender, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                  color    = "white",
                  size     = 4,
                  fontface = 'bold',
                  hjust    = 1.3,
                  vjust    = 0.4) +
        theme_minimal() +
        theme(legend.position = 'none')

ed <- ggplot(cat_feat %>% 
                group_by(education_level, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
             aes(education_level, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                    color    = "white",
                    size     = 4,
                    fontface = 'bold',
                    hjust    = 1.3,
                    vjust    = 0.4) +
        theme_minimal() +
        theme(axis.title.x = element_blank(),
              legend.position = 'none')

inc <- ggplot(cat_feat %>% 
                group_by(income_category, churned) %>% 
                summarise(n = n()) %>% 
                mutate(percent_churned = prop.table(n)*100) %>% 
                filter(churned == 'yes'),
              aes(income_category, percent_churned)) +
        geom_col() +
        coord_flip() +
        geom_text(aes(label = round(percent_churned, digits = 1)),
                    color    = "white",
                    size     = 4,
                    fontface = 'bold',
                    hjust    = 1.3,
                    vjust    = 0.4) +
        theme_minimal() +
        theme(axis.title.x = element_blank(),
              legend.position = 'none')

ed + inc + gen +
    plot_layout(ncol = 2, heights = c(2,0.7)) +
    plot_annotation(title = "Percentage of Each Categorical Variable that Churned")
```

```{r chunk12, cache=TRUE}
cat_feat %>%
  pivot_longer(gender:income_category) %>%
  ggplot(aes(y = value, fill = churned)) +
  geom_bar(position = "fill") +
  facet_wrap(vars(name), scales = "free", ncol = 1) +
  theme_minimal() +
  labs(x = NULL, y = NULL)
```

At least the data has consistantly low churn (12%-21%) across all categorical values.
I'd guess the numeric features are more important based on what we've seen.

## 🧮 AWS Setup {#aws}

Now it's time to model. You need to set up your own AWS account and get RStudio
connected. Alex Lemm has a great repo that guides you through configuring RStudio to
work with AWS tools
[github.com/alex23lemm](https://github.com/alex23lemm/AWS-SageMaker-Fundamentals-R-Workshop).
You'll find a more detailed explaination of my code chunks within his projects if you
want a deeper understanding.

We need to set up our AWS SageMaker session using a conda or miniconda environment
and the `{reticulate}` package.

This code works on the AWS 'Free Tier' btw.

```{r chunk13, eval=FALSE}
use_condaenv("sagemaker-r", required = TRUE)
sagemaker <- import("sagemaker")
session <- sagemaker$Session()
```

Select the S3 bucket you want to upload your data and models to. Then create a `data`
and `models` folder. The `data` folder will store training, validation, test, and
holdout data XGBoost will use for this project. The `models` folder will store every
training job we run in this project, with an `output` subfolder containing the model
artifacts. Any model used to make predictions will also have a `predictions`
subfolder.

```{r chunk14, eval=FALSE}
# bucket <- session$default_bucket()
bucket <- "twarczak-sagemaker3"
project <- "churn"    # keep short. 32 character max job_name
data_path <- paste0("s3://", bucket, "/", project, "/", "data")
models_path <- paste0("s3://", bucket, "/", project, "/", "models")
```

### Data splitting

Split the `churn_data` into a train/validation/test. We will use the SageMaker
built-in XGBoost algorithm and Bayesian Optimation for hyperparameter tuning. Info
about this version of XGBoost, hyperparameters, metrics, etc. can be found here
[AWS_XGBoost](https://docs.aws.amazon.com/sagemaker/latest/dg/xgboost.html).

I am also using the original data, not the dataframes I manipulated for our EDA.

```{r chunk15, cache=TRUE, eval=FALSE}
# Make 70/15/15 Train/Validate/Test split
set.seed(333)
splits  <- initial_split(churn_data, prop = 0.70, strata = attrition_flag)

train <- training(splits)
other  <- testing(splits)

splits2 <- initial_split(other, prop = 0.50, strata = attrition_flag)
validation <- training(splits2)
test <- testing(splits2)

print(splits)
print(splits2)
```

### Preprocessing the data

Let's make a simple recipe for our data. XGBoost doesn't need numeric variable
transformations, but we need to one-hot encode the categorical features. SageMaker
XGBoost doesn't want id variables and the target (`attrition_flag`) must be the first
column. We'll also remove the table header later.

```{r chunk16, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
churn_rec <- recipe(attrition_flag ~ ., data = train) %>% 
             step_dummy(all_nominal_predictors(), one_hot = TRUE) %>% 
             step_rm(id) %>% 
             prep()

churn_training <- churn_rec %>% 
                  juice() %>% 
                  select(attrition_flag, everything()) # just to put attrition_flag as 1st column

churn_validation <- bake(churn_rec, new_data = validation) %>% 
                    select(attrition_flag, everything()) # just to put attrition_flag as 1st column

churn_test <- bake(churn_rec, new_data = test) %>% 
              select(attrition_flag, everything()) # just to put attrition_flag as 1st column

id_holdout <- holdout %>% select(id) 
churn_holdout <- bake(churn_rec, new_data = holdout)
```

Great. Now all our data is transformed and ready for SageMaker XGBoost. First though,
let's run the training data through a basic XGBoost algo in R, just to look at
feature importance.

```{r chunk17, message=FALSE, warning=FALSE, cache=TRUE}
bst <- xgboost::xgboost(data = data.matrix(subset(churn_training, select = -attrition_flag)), 
               label = churn_training$attrition_flag, max_depth = 2, eta = 0.5, nthread = 2, nrounds = 100, 
               objective = "binary:logistic", verbose = FALSE)

imp_bst <- xgb.importance(model = bst)
xgb.ggplot.importance(imp_bst)
```

```{r chunk17.5, include=FALSE}
vip::vip(bst, geom = 'point', num_features = 15)
```

We should have looked harder at `total_revolving_balance`. It might be more important
than we thought, but otherwise, this mostly agrees with our EDA assessment. We could
go back and remove the low importance features, like `income_category`,
`education_level`, and `gender`, but not this time.

Save the pre-processed data. The test set must lack the target variable.

Save `churn_holdout_id` to join to the predicted values at the very end for our
SLICED submission.

```{r chunk18, message=FALSE, eval=FALSE}

dir.create("../2021-08-01-sagemaker-r-xgb-churn/data") # local

write_csv(churn_training, 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_training.csv", 
          col_names = FALSE)
write_csv(churn_validation, 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_validation.csv", 
          col_names = FALSE)
write_csv(churn_test %>% select(-attrition_flag), 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_test.csv", 
          col_names = FALSE)
write_csv(churn_test, 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_test_2.csv", 
          col_names = TRUE) # Need this later to use with results


write_csv(churn_holdout, 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_holdout.csv", 
          col_names = FALSE) 
write_csv(id_holdout, 
          "../2021-08-01-sagemaker-r-xgb-churn/data/churn_holdout_id.csv",
          col_names = FALSE) # Need this later to use with holdout results
```

`upload()` from `S3Uploader` moves the preprocessed data to the S3 bucket.

```{r chunk19, cache=TRUE, eval=FALSE}
s3_uploader <- sagemaker$s3$S3Uploader()

s3_train <- s3_uploader$upload(local_path = "../2021-08-01-sagemaker-r-xgb-churn/data/churn_training.csv", 
                               desired_s3_uri = data_path)

s3_validation <- s3_uploader$upload(local_path = "../2021-08-01-sagemaker-r-xgb-churn/data/churn_validation.csv",
                                    desired_s3_uri = data_path)

s3_test <- s3_uploader$upload(local_path = "../2021-08-01-sagemaker-r-xgb-churn/data/churn_test.csv", 
                              desired_s3_uri = data_path)
```

## 🏗 Train XGBoost model {#build}

### Step 1 - Create an Estimator object

Need

-   `image_name` - Location of built-in XGBoost docker container image in AWS Elastic
    Container Registry (ECR)

-   `role` - AWS Identify and Access Management (IAM)

-   `instance_count` - Number of EC2 instances

-   `instance_type` - Choice of instance

-   `volume_size` - Size in GB of Amazon Elastic Block Store (EBS)

-   `output_path` - Path to S3 bucket for storing training results

-   `sagemaker_session` - Session object that manages interactions with SageMaker
    APIs

Setting a `max_run`, `use_spot_instances`, and `max_wait` is optional, but will save
you 70-90% on your compute purchase (in my experience).

Note: With `{reticulate}`, integers in R need to be converted to "long" objects in
Python, hence the "L" after integers in the script.

```{r chunk20, cache=TRUE, eval=FALSE}
region <- session$boto_region_name

container <- sagemaker$image_uris$retrieve(framework = "xgboost", 
                                           region    = region, 
                                           version   = "1.3-1" )

role_arn <- Sys.getenv("SAGEMAKER_ROLE_ARN")

xgb_estimator <- sagemaker$estimator$Estimator(image_uri          = container,
                                               role               = role_arn,
                                               instance_count     = 1L,
                                               instance_type      = "ml.m5.4xlarge",  
                                               volume_size        = 10L, 
                                               max_run            = 300L,
                                               output_path        = models_path,
                                               sagemaker_session  = session,
                                               use_spot_instances = TRUE,
                                               max_wait           = 300L )
```

### Step 2 - Set static hyperparameters

No tuning yet. Full list of SageMaker built-in XGBoost hyperparameters here:
[XGBoost](https://docs.aws.amazon.com/sagemaker/latest/dg/xgboost_hyperparameters.html).

Use AUC as metric

```{r chunk21, cache=TRUE, eval=FALSE}
xgb_estimator$set_hyperparameters(objective        = "binary:logistic",
                                  eval_metric      = "auc",
                                  max_depth        = 5L, 
                                  eta              = 0.1,
                                  num_round        = 100L,
                                  colsample_bytree = 0.4,
                                  alpha            = 10L,
                                  min_child_weight = 1.1,
                                  subsample        = 0.7)
```

### Step 3 - Define S3 location of data sets and training job name

```{r chunk22, cache=TRUE, eval=FALSE}
# Create training job name based project organization principles
algo <- "xgb"    # keep short. 32 character max job_name
timestamp <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S") # timestamp
job_name_x <- paste(project, algo, timestamp, sep = "-")
s3_train_input <- sagemaker$inputs$TrainingInput(s3_data = s3_train,
                                                 content_type = 'csv')
s3_valid_input <- sagemaker$inputs$TrainingInput(s3_data = s3_validation,
                                                 content_type = 'csv')
input_data <- list('train'      = s3_train_input,
                   'validation' = s3_valid_input)
```

### Step 4 - Start training job

```{r chunk23, cache=TRUE, eval=FALSE}
xgb_estimator$fit(inputs = input_data,
                  job_name = job_name_x,
                  wait = FALSE)  # If TRUE, call will wait until job completes
```

```{r chunk24, cache=TRUE}
session$describe_training_job(job_name_x)[["TrainingJobStatus"]]
```

### Step 5 - Evaluate training results

```{r chunk25, cache=TRUE, eval=FALSE}
training_job_stats <- session$describe_training_job(job_name = job_name_x)
final_metrics_1 <-  map_df(training_job_stats$FinalMetricDataList, 
                           ~tibble(metric_name = .x[["MetricName"]],
                                   value = .x[["Value"]]))
```

```{r chunk25.2, cache=TRUE}
final_metrics_1
```


### Step 6 - Fit model on test data

Use Batch Transform, a SageMaker feature for generating batch inferences.

Create Transformer object

-   `strategy` - 'MultiRecord' or 'SingleRecord'

-   `assemble_with` - 'Line' places each prediction on its own line

-   `output_path` - S3 location for storing prediction results

```{r chunk26, cache=TRUE, eval=FALSE}
predictions_path_1 <- paste0(models_path, "/", job_name_x, "/predictions") 

xgb_batch_predictor_1 <- xgb_estimator$transformer(instance_count     = 1L, 
                                                   instance_type      = "ml.m5.4xlarge", 
                                                   strategy           = "MultiRecord",
                                                   assemble_with      = "Line",
                                                   output_path        = predictions_path_1 )
```

Fit model to test data with `transform()`

-   `split_type` - Must use "Line" with csv files

```{r chunk27, cache=TRUE, eval=FALSE}
xgb_batch_predictor_1$transform(data         = s3_test, 
                                content_type = 'text/csv',
                                split_type   = "Line",
                                job_name     = job_name_x,
                                wait         = FALSE) # If TRUE, call waits until job completes
```

```{r chunk28, cache=TRUE}
session$describe_transform_job(job_name_x)[["TransformJobStatus"]]
```

### Step 7 - Download & evaluate predictions

```{r chunk29, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
s3_downloader <- sagemaker$s3$S3Downloader()
s3_test_predictions_path_1 <- s3_downloader$list(predictions_path_1)
 
dir.create("./predictions")
s3_downloader$download(s3_test_predictions_path_1, "./predictions")
 
test_predictions_1 <- read_csv("./predictions/churn_test.csv.out",
                               col_names = FALSE) %>% 
                      pull(X1)

test_results_1 <- tibble(truth       = churn_test$attrition_flag,
                         predictions = test_predictions_1 )
```

````{r chunk29.2, cache=TRUE}
head(test_results_1)
```

```{r chunk30, message=FALSE, warning=FALSE, cache=TRUE}
roc_obj_1 <- pROC::roc(test_results_1$truth,
                       test_results_1$predictions,
                       plot        = TRUE,         
                       grid        = TRUE,
                       print.auc   = TRUE,
                       legacy.axes = TRUE, 
                       main        = "ROC curve for XGBoost classification",
                       show.thres  = TRUE,
                       col         = "red2" )
```

AUC: 0.989

```{r chunk31, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
conf_matrix_1 <- caret::confusionMatrix(factor(ifelse(test_results_1$predictions >= 0.5, 1, 0),
                                                levels = c("0", "1"), 
                                                labels = c("retained", "churned")),
                                         factor(test_results_1$truth, 
                                                levels = c(0, 1), 
                                                labels = c("retained", "churned")),
                                         positive = "churned")
```
```{r chunk31.2, cache=TRUE}
conf_matrix_1
```

-   Accuracy: **0.9671**

-   Sensitivity: **0.8294**

-   Specificity: **0.9933**

-   Kappa: **0.8704**

-   False Negatives: **29**

-   False Positives: **6**

## 🔧 Tune XGBoost {#tune}

### Step 8 - Set hyperparameters

**Static**

```{r chunk32, cache=TRUE, eval=FALSE}
xgb_estimator$set_hyperparameters(objective        = "binary:logistic",
                                  min_child_weight = 1 )
```

**Tunable**

```{r chunk33, cache=TRUE, eval=FALSE}
hyperp_ranges <- list(num_round = sagemaker$tuner$IntegerParameter(50L, 500L),
                      max_depth = sagemaker$tuner$IntegerParameter(2L, 15L),
                      eta =sagemaker$tuner$ContinuousParameter(0.01,0.4,"Logarithmic"),
                      colsample_bytree = sagemaker$tuner$ContinuousParameter(0.3,0.8),
                      alpha = sagemaker$tuner$IntegerParameter(1L, 15L))
```

Create `HyperparameterTuner` object.

SageMaker supports two tuning methods: Bayesian optimization (default) and random
search. There is currently no grid search option. Let's stick with Bayesian.

-   `max_jobs` - Maximum number of training jobs executed

-   `max_parallel_job`: Number of training jobs executed in parallel

```{r chunk34, cache=TRUE, eval=FALSE}
tuner <- sagemaker$tuner$HyperparameterTuner(estimator             = xgb_estimator,
                                             objective_metric_name = "validation:auc",
                                             objective_type        = "Maximize",
                                             hyperparameter_ranges = hyperp_ranges, 
                                             strategy              = "Bayesian",
                                             max_jobs              = 100L,
                                             max_parallel_jobs     = 3L )
```

Define S3 location of data sets and tuning job name.

```{r chunk35, cache=TRUE, eval=FALSE}
algo <- "xgb"
timestamp <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
job_name_y <- paste(project, algo, timestamp, sep = "-")
s3_train_input <- sagemaker$inputs$TrainingInput(s3_data = s3_train,
                                                 content_type = 'csv')
s3_valid_input <- sagemaker$inputs$TrainingInput(s3_data = s3_validation,
                                                 content_type = 'csv')
input_data <- list('train'      = s3_train_input,
                   'validation' = s3_valid_input)
```

### Step 9 - Start tuning job

```{r chunk36, cache=TRUE, eval=FALSE}
tuner$fit(inputs   = input_data, 
          job_name = job_name_y,
          wait     = FALSE ) # If TRUE, call will wait until job completes
```

```{r chunk37, cache=TRUE}
session$describe_tuning_job(job_name_y)[["HyperParameterTuningJobStatus"]]
```

### Step 10 - Evaluate tuning job results

```{r chunk38, cache=TRUE, eval=FALSE}
tuning_job_results <- sagemaker$HyperparameterTuningJobAnalytics(job_name_y)
tuning_results_df <- tuning_job_results$dataframe()
```
```{r chunk38.2, cache=TRUE}
head(tuning_results_df)
```

Plot a time series chart that shows how AUC developed over 50 training jobs, tuned by
the Bayesian optimizer.

```{r chunk39, fig.align='center', message=FALSE, warning=FALSE, cache=TRUE}
ggplot(tuning_results_df, aes(TrainingEndTime, FinalObjectiveValue)) +
  geom_point() +
  xlab("Time") +
  ylab(tuning_job_results$description()$TrainingJobDefinition$StaticHyperParameters$`_tuning_objective_metric`) +
  ggtitle("Hyperparameter tuning AUC",  
          "Progression over the period of all 100 training jobs") +
  theme_minimal()
```

The next plots show how the Bayesian optimizer focused on the region of the search
space that produced the best models.

```{r chunk40, fig.align='center', cache=TRUE}
ggplot(tuning_results_df, aes(num_round, eta)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_viridis("AUC", option = "H") +
  ggtitle("Hyperparameter tuning AUC", 
          "Using a Bayesian strategy") +
  theme_dark()
```

```{r chunk40.2, fig.align='center', cache=TRUE}
ggplot(tuning_results_df, aes(colsample_bytree, alpha)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_viridis("AUC", option = "H") +
  ggtitle("Hyperparameter tuning AUC", 
          "Using a Bayesian strategy") +
  theme_dark()
```

```{r chunk40.3, fig.align='center', cache=TRUE}
ggplot(tuning_results_df, aes(num_round, max_depth)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_viridis("AUC", option = "H") +
  ggtitle("Hyperparameter tuning AUC", 
          "Using a Bayesian strategy") +
  theme_dark()
```

## 🏆 Select Best Models {#best}

```{r chunk41, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}

best_tuned_model <- tuning_results_df %>%
                    #filter(FinalObjectiveValue == max(FinalObjectiveValue)) %>%
                    arrange(desc(FinalObjectiveValue)) %>% 
                    head(n=1) %>% pull(TrainingJobName)

#best_tuned_model
training_job_stats <- session$describe_training_job(job_name = best_tuned_model)

final_metrics_2 <-  map_df(training_job_stats$FinalMetricDataList, 
                           ~tibble(metric_name = .x[["MetricName"]],
                                   value       = .x[["Value"]]))
```
```{r chunk41.2, cache=TRUE}
final_metrics_2
```

### Step 11 - Create transformer object

```{r chunk42, cache=TRUE, eval=FALSE}
predictions_path_2 <- paste0(models_path, "/", best_tuned_model, "/predictions")

session$create_model_from_job(best_tuned_model)

xgb_batch_predictor_2 <- sagemaker$transformer$Transformer(model_name     = best_tuned_model,
                                                           instance_count = 1L, 
                                                           instance_type  = "ml.m5.4xlarge",   
                                                           strategy       = "MultiRecord",  
                                                           assemble_with  = "Line",
                                                           output_path    = predictions_path_2)
```
```{r chunk42.2, cache=TRUE}
best_tuned_model
```


### Step 12 - Start batch prediction job

```{r chunk43, cache=TRUE, eval=FALSE}
xgb_batch_predictor_2$transform(data         = s3_test, 
                                content_type = 'text/csv',
                                split_type   = "Line",
                                job_name     = best_tuned_model,
                                wait         = FALSE) # If TRUE, call will wait until job completes

```

```{r chunk44, cache=TRUE}
session$describe_transform_job(best_tuned_model)[["TransformJobStatus"]]
```

### Step 13 - Download test set predictions

```{r chunk45, cache=TRUE, eval=FALSE}
s3_downloader <- sagemaker$s3$S3Downloader()

s3_test_predictions_path_2 <- s3_downloader$list(predictions_path_2)
 
dir.create("./predictions")

s3_downloader$download(s3_test_predictions_path_2, "./predictions")
 
test_predictions_2 <- read_csv("./predictions/churn_test.csv.out",
                               col_names = FALSE) %>% 
                      pull(X1)

churn_test <- read_csv("./data/churn_test_2.csv")

test_results_2 <- tibble(truth       = churn_test$attrition_flag,
                         predictions = test_predictions_2)
```

```{r chunk45.2, cache=TRUE}
head(test_results_2)
```

### Step 14 - Evaluate test set predictions

```{r chunk46, message=FALSE, warning=FALSE, cache=TRUE}
roc_obj_2 <- roc(test_results_2$truth, 
                 test_results_2$predictions,
                 plot        = TRUE,         
                 grid        = TRUE,
                 print.auc   = TRUE,
                 legacy.axes = TRUE, 
                 main        = "ROC curve for XGBoost classification using the best model",
                 show.thres  = TRUE,
                 col         = "red2" )
```

Confusion matrix

```{r chunk47, cache=TRUE, eval=FALSE}
conf_matrix_2 <- confusionMatrix(factor(ifelse(test_results_2$predictions >= 0.5, 1, 0),
                                        levels = c("0", "1"), 
                                        labels = c("retained", "churned")),
                                 factor(test_results_2$truth, 
                                        levels = c(0, 1), 
                                        labels = c("retained", "churned")),           
                                 positive = "churned")
```

```{r chunk47.2, cache=TRUE}
conf_matrix_2
```

-   Accuracy: **0.9756**

-   Sensitivity: **0.8706**

-   Specificity: **0.9955**

-   Kappa: **0.9049**

-   False Negatives: **22**

-   False Positives: **4**

```{r chunk48, eval=FALSE, cache=FALSE, include=FALSE}
test_results_pred <- test_results_2 %>% 
                     mutate(pred = factor(ifelse(predictions >= 0.5, 1, 0), 
                                          levels = c("0", "1")))

gmodels::CrossTable(x          = test_results_pred$pred, 
                    y          = test_results_pred$truth,
                    prop.chisq = FALSE, 
                    prop.r     = FALSE, 
                    prop.c     = FALSE, 
                    prop.t     = FALSE,
                    dnn        = c('predicted churn', 'actual churn'))
```

Save this model as our `churn_model`. `boto_client$list_models()` returns the model
with the best 'Objective metric value' as the first object.

```{r chunk49, cache=TRUE, eval=FALSE}
boto_client <- session$boto_session$client("sagemaker")
churn_model <- boto_client$list_models()[["Models"]] %>% 
  map_chr("ModelName") %>% 
  .[[1]]
```
```{r chunk49.2, cache=TRUE}
churn_model
```

### Step 15 - Create endpoint configuration

We can deploy the model so we (or anyone we give permission to) can use it to make
more bank customer churn predictions. We still need to run inference on our
`churn_holdout` data and submit those predictions to the SLICED competition for
scoring.

```{r chunk50, results='hide', cache=TRUE, eval=FALSE}
config_name <- paste0(churn_model, "-config")
session$create_endpoint_config(name = config_name,
                               model_name =  churn_model, 
                               initial_instance_count = 1L, 
                               instance_type = "ml.m5.4xlarge")
```

### Step 16 - Create endpoint

```{r chunk51, results='hide', cache=TRUE, eval=FALSE}
endpoint_name <- "churn-endpoint"
session$create_endpoint(endpoint_name = endpoint_name, 
                        config_name = config_name,
                        wait = FALSE)
```

Check via the API when the endpoint is successfully deployed and available. Might
take 5-10 minutes.

```{r chunk52, cache=TRUE}
boto_client$describe_endpoint(EndpointName = endpoint_name)[["EndpointStatus"]]
```

### Step 17 - Make real-time predictions against the endpoint (not batch)

-   Predictor object - Makes prediction requests to the inference endpoint
-   Serializer object - Encodes data for inference endpoint
-   Deserializer object - Decodes data from inference endpoint

```{r chunk53, cache=TRUE, eval=FALSE}
csv_serializer <- sagemaker$serializers$CSVSerializer()
csv_deserializer <- sagemaker$deserializers$CSVDeserializer()
churn_predictor <- sagemaker$predictor$Predictor(endpoint_name     = endpoint_name, 
                                                 sagemaker_session = session, 
                                                 serializer        = csv_serializer,
                                                 deserializer      = csv_deserializer)
```

Let's just check the first 5 test_set entries to make sure it worked. We already ran
a batch transform job, so the results should be the same. This is just to test the
endpoint.

```{r chunk54, message=FALSE, cache=TRUE, eval=FALSE}
test_set <- read_csv("./data/churn_test.csv", 
                     col_names = FALSE, 
                     n_max     = 5) %>% 
            as.matrix()
real_time_predictions_1 <- churn_predictor$predict(data = test_set) %>% 
                           .[[1]] %>% 
                           as.numeric()
```

```{r chunk55, message=FALSE, cache=TRUE, eval=FALSE}
batch_predictions <- read_csv("./predictions/churn_test.csv.out", 
                              col_names = FALSE, n_max = 5) %>% 
                     .[[1]]
```  
```{r chunk55.2, cache=TRUE}
data.frame(real_time_predictions_1, batch_predictions)
```

Perfect!

1 model, 2 prediction pipelines, same results.

## 🎯 Predict w/ Holdout {#holdout}

Use the endpoint to predict churn from our holdout data.

```{r chunk56, message=FALSE, cache=TRUE, eval=FALSE}
holdout_set <- read_csv("./data/churn_holdout.csv", 
                        col_names = FALSE) %>% 
               as.matrix()
real_time_predictions_2 <- churn_predictor$predict(data = holdout_set) %>% 
                           .[[1]] %>% 
                           as.numeric()
```

## 📬 Submission {#submit}

Save and submit to SLICED s01e07 on Kaggle.

```{r chunk57, message=FALSE, warning=FALSE, cache=TRUE, eval=FALSE}
churn_holdout_id <- read_csv("data/churn_holdout_id.csv", col_names = FALSE)
submission <- data.frame(churn_holdout_id, real_time_predictions_2) %>% 
              rename(id = X1, attrition_flag = real_time_predictions_2)
write.csv(submission, "sliced_s01_e07_submission.csv", row.names = FALSE)
```

The evaluation algorithm for this competition was **LogLoss**, so the lowest score
wins. My submission scored **0.07622**, which would have been 8th place (of 36
competitors). The winning score was **0.06800**. Competitors only had 2 hours to
submit, so this is not a fair comparison, but it does show how a well-tuned XGBoost
model can make great predictions without much feature engineering. To improve the
model, we could log transform certain numeric variables, up/down-sample with
`{themis}` to balance the data, remove low-performing features, and/or tune
additional hyperparameters. There should be room to improve model performance.

(Delete the endpoint so your not billed more \$\$\$)

```{r chunk58, cache=TRUE}
session$delete_endpoint(endpoint_name)
```
