---
title: "Multi-class Prediction of Austin housing prices with XGBoost and AWS SageMaker in R"
author: Todd Warczak
date: '2021-08-19'
slug: XGB-Multi-Class
categories:
  - rstats
  - AWS
  - SageMaker
  - XGBoost
  - R
  - SHAP
  - Geospacial 
tags:
  - reticulate
  - AWS
  - SageMaker
  - XGBoost
  - multiclass
  - SHAP
  - R
  - rstats
  - NLP
  - ggmaps
  - Leaflet
  - Geospacial
subtitle: 'Interpretable XGBoost with SHAP'
summary: "In this post I explore an Austin Housing dataset and predict binned housing price. EDA includes static and interactive geospacial feature maps and feature engineering using natural language processing (NLP). In my RStudio I utilize a conda environment to access SageMaker's XGBoost, S3, and EC2 for distributed computing. After model training/tuning/evaluation I run SHapley Additive exPlanations (SHAP) to understand what the XGBoost model considered important features and how it clustered similar observations. I run batch inference with the best performing model to predict the price of Austin houses in holdout data. I then submitted predictions to the Kaggle competition. Multiclass Log Loss was the evaluation algorithm and my submission scored ___, which would have placed __ in the live competition."
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: true
---

```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE, 
                      echo = TRUE, dpi = 300, cache.lazy = FALSE,
                      tidy = "styler", fig.width = 8, fig.height = 5)
library(scales)
library(tidyverse)
library(tidymodels)
library(ggmap)
library(leaflet)
library(scico)
library(patchwork)
library(tidytext)
library(paws)
library(reticulate)    # calling the SageMaker Python SDK from R
theme_set(ggdark::dark_theme_minimal())


#library(finetune)
```

In this post I explore an Austin, TX real estate dataset and predict binned housing price. EDA includes static and interactive geospacial feature maps and feature engineering using natural language processing (NLP). In my RStudio I utilize a conda environment to access SageMaker's XGBoost, S3, and EC2 for distributed computing. After model training/tuning/evaluation I run SHapley Additive exPlanations (SHAP) to understand what the XGBoost model considered important features and how it clustered similar observations. I ran batch inference with the best performing model to predict the price of Austin houses in holdout data and then submitted predictions to the [Kaggle](https://www.kaggle.com/c/sliced-s01e11-semifinals/) competition. Multiclass Log Loss was the evaluation algorithm and my submission scored ___ (lower is better), which would have placed __ in the live competition.

## EDA

Our modeling goal is to predict  given features about the real
estate listing. This is a multiclass classification challenge, where . The main data set
provided is in a CSV file called `training.csv`.

```{r}
train_raw <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-austin-r-xgb-house-price/train.csv")

holdout <- read_csv("~/Documents/R/data_warz/content/post/2021-08-01-austin-r-xgb-house-price/test.csv")
names(train_raw)
names(holdout) # lacks priceRange
table(train_raw$priceRange)
```

You can watch [this week's full episode of
SLICED](https://www.twitch.tv/videos/1114553508) to see lots of exploratory data
analysis and visualization of this dataset, but let's just make a few data
visualization for context in this blog post.

How is price distributed across Austin?

```{r fig.height=4, fig.width=4, warning=FALSE, message=FALSE}
qmplot(longitude, latitude, data = train_raw, 
       color = priceRange, darken = .4, alpha = I(.6))
```

```{r fig.height=8, fig.width=8}
q <- train_raw %>%
     mutate(priceRange_bin = parse_number(priceRange))
#table(q$priceRange)
#scico::scico_palette_show(palettes = scico_palette_names())

pal <- colorNumeric(palette = scico(6, palette = 'buda'),
                    domain = q$priceRange_bin)

m <- leaflet(q) %>% 
     addTiles() %>% 
     addCircles(lng = ~longitude, lat = ~latitude, 
                color = ~pal(priceRange_bin), opacity = 0.8) %>% 
     setView(lng = -97.75, lat = 30.3, zoom = 12 ) %>% 
     addProviderTiles("CartoDB.DarkMatterNoLabels")%>%
     addProviderTiles("Stamen.TonerLines") %>%
     addProviderTiles("Stamen.TonerLabels") %>% 
     addLegend("bottomright", 
               pal       = pal, 
               values    = ~priceRange_bin,
               title     = "Price", 
               labFormat = labelFormat(prefix = "$"),
               opacity   = 1)
m
```

Let's look at this distribution and compare it to some other variables available in
the dataset. We can create a little plotting function [using
`{{}}`](https://dplyr.tidyverse.org/articles/programming.html#indirection) to quickly
iterate through, and put them together with
[patchwork](https://patchwork.data-imaginist.com/).

```{r, fig.width=10, fig.height=14}
plot_hex <- function(var, title) {
  q %>%
    ggplot(aes(longitude, latitude, z = {{var}})) +
    stat_summary_hex(bins = 40) +
    scico::scale_fill_scico(palette = 'imola') +
    labs(fill = "Avg", title = title) +
    ggdark::dark_theme_minimal()  
}

# names(q)

(plot_hex(priceRange_bin,"Price") + plot_hex(avgSchoolRating, "School rating")) /
(plot_hex(yearBuilt,"Year built") + plot_hex(log(lotSizeSqFt),"Lot size (log)"))/
(plot_hex(garageSpaces,"Garages") + plot_hex(MedianStudentsPerTeacher, "Median Student/Teacher")) / 
(plot_hex(numOfBathrooms, "Bathrooms") + plot_hex(numOfBedrooms,"Bedrooms")) 
```

```{r}
plot_hex(priceRange_bin,"Price by city") +
  facet_wrap(~city)

plot_hex(priceRange_bin,"Price by hasSpa") +
  facet_wrap(~hasSpa)

plot_hex(numOfPatioAndPorchFeatures, "# Patio Features")
```

Notice the east/west gradients as well as the radial changes. I went to grad school
in Austin and this all look very familiar to me!

```{r fig.width=10, fig.height=10}
dense <- q %>%
         select(-c(uid, city, description, homeType, 
                   hasSpa, priceRange_bin, yearBuilt)) %>%
         pivot_longer(cols = c(garageSpaces, avgSchoolRating, MedianStudentsPerTeacher,
                               numOfBathrooms, numOfBedrooms, longitude, latitude), 
                      names_to = "feature", 
                      values_to = "value")

ggplot(dense, aes(value, fill = priceRange)) +
  geom_density(alpha = .5) +
  facet_wrap(~ feature, scales = "free") +
  labs(title = "Numeric features impacting priceRange") +
  ggdark::dark_theme_minimal() +
  theme(legend.position = c(0.45, 0.2),
        legend.background = element_rect(fill = "black", color = "white"))
```

## Finding words related to price

The `description` variable contains text from each real estate listing. We could try
to use the text features directly in modeling, [as described in our
book](https://smltar.com/), but I've found that often isn't great for boosted tree
models (which tend to be what works best overall in an environment like SLICED).
Let's walk through another option which may work better in some situations, which is
to use some separate analysis to identify important words and then create dummy
variables indicating whether any given listing has those words.

Let's start by tidying the `description` text.

```{r}
austin_tidy <- 
  train_raw %>%
  mutate(priceRange = parse_number(priceRange) + 100000) %>%
  unnest_tokens(word, description) %>%
  anti_join(get_stopwords()) # remove stop words

austin_tidy %>%
  count(word, sort = TRUE)
```

Next, let's compute word frequencies per price range for the top 100 words.

```{r}
#dim(q)
top_words <- 
  austin_tidy %>%
  count(word, sort = TRUE) %>%
  filter(!word %in% as.character(1:5)) %>% # removing numbers
  slice_max(n, n = 100) %>% 
  pull(word)

word_freqs <- 
  austin_tidy %>%
  count(word, priceRange) %>%
  complete(word, priceRange, fill = list(n = 0)) %>%
  group_by(priceRange) %>%
  mutate(price_total = sum(n),
         proportion = n / price_total) %>%
  ungroup() %>%
  filter(word %in% top_words)

word_freqs
```

Now let's use modeling to find the words that are **increasing** with price and those
that are **decreasing** with price.

```{r}
word_mods <-
  word_freqs %>%
  nest(data = c(priceRange, n, price_total, proportion)) %>%
  mutate(model = map(data, ~ glm(cbind(n, price_total) ~ priceRange, ., 
                                 family = "binomial")),
         model = map(model, tidy)) %>%
  unnest(model) %>%
  filter(term == "priceRange") %>%
  mutate(p.value = p.adjust(p.value)) %>%
  arrange(-estimate)

word_mods
```

These are the words that we'd like to try to detect and use in feature engineering
for our xgboost model, rather than using all the text tokens as features
individually.

```{r}
higher_words <-
  word_mods %>%
  filter(p.value < 0.05) %>%
  slice_max(estimate, n = 12) %>%
  pull(word)

lower_words <- 
  word_mods %>%
  filter(p.value < 0.05) %>%
  slice_max(-estimate, n = 12) %>%
  pull(word)
```

We can look at these changes with price directly. For example, these are the words
most associated with price decrease.

```{r, fig.width=12, fig.height=14}
high <- word_freqs %>% 
  filter(word %in% higher_words) %>%
  ggplot(aes(priceRange, proportion, color = word)) +
  geom_line(size = 2.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(vars(word), scales = "free") +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = NULL, y = "% of word in home description at price",
       title = "12 Expensive Words") +
  ggdark::dark_theme_minimal() 

low <- word_freqs %>% 
  filter(word %in% lower_words) %>%
  ggplot(aes(priceRange, proportion, color = word)) +
  geom_line(size = 2.5, alpha = 0.7, show.legend = FALSE) +
  facet_wrap(vars(word), scales = "free") +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = NULL, y = "% of word in home description at price",
       title = "12 Cheap Words") +
  ggdark::dark_theme_minimal() 

high/low
```

Cheaper houses are "great" but not expensive houses, and apparently you don't need to
mention the location ("close", "minutes", "location") of more expensive houses.

## Build a model

Let's start our modeling by setting up our "data budget", as well as the metrics
(this challenge was evaluate on multiclass log loss).

```{r}
train_raw$priceRange <- factor(recode(train_raw$priceRange,
                                      "0-250000"      = "0", 
                                      "250000-350000" = "1",
                                      "350000-450000" = "2",
                                      "450000-650000" = "3", 
                                      "650000+"       = "4"), 
                               levels = c("0", "1", "2", "3", "4"))

head(train_raw$priceRange, n = 20)
```


```{r}
# Make 75/12.5/12.5 Train/Validate/Test split
set.seed(333)
splits  <- initial_split(train_raw, prop = 0.75, strata = priceRange)

train <- training(splits)
other  <- testing(splits)

splits2 <- initial_split(other, prop = 0.50, strata = priceRange)
validation <- training(splits2)
test <- testing(splits2)

print(splits)
print(splits2)
```

For feature engineering, let's use basically everything in the dataset (aside from
`city`, which was not a very useful variable) and [create dummy or indicator
variables using
`step_regex()`](https://recipes.tidymodels.org/reference/step_regex.html). The idea
here is that we will detect whether these words associated with low/high price are
there and create a yes/no variable indicating their presence or absence.

```{r}
higher_words_6 <- word_mods %>%
                   filter(p.value < 0.05) %>%
                   slice_max(estimate, n = 6) %>%
                   pull(word)

lower_words_5 <- word_mods %>%
                  filter(p.value < 0.05) %>%
                  slice_max(-estimate, n = 5) %>%
                  pull(word)

higher_pat <- glue::glue_collapse(higher_words_6, sep = "|")
lower_pat <- glue::glue_collapse(lower_words_5, sep = "|")

rec_austin <-
  recipe(priceRange ~ ., data = train) %>%
  step_regex(description, pattern = higher_pat, result = "high_price_words") %>%
  step_regex(description, pattern = lower_pat, result = "low_price_words") %>%
  step_rm(c(uid, city, description, hasSpa, numOfPatioAndPorchFeatures)) %>%
  step_novel(homeType) %>% 
  # assign novel factor level if new data has homeType category unseen in train data
  step_unknown(homeType) %>% 
  # assign missing value to level "unknown"
  step_other(homeType, threshold = 0.02) %>%
  # low occuring values get pooled into "other" categoory
  step_dummy(all_nominal_predictors()) %>%
  #step_nzv(all_predictors()) %>% 
  # removes highly sparse and unbalanced predictors
  prep()
```

```{r}
austin_training <- rec_austin %>% 
                  juice() %>% 
                  select(priceRange, everything()) # put priceRange as 1st column
head(austin_training)
#table(austin_training$priceRange)

austin_validation <- bake(rec_austin, new_data = validation) %>% 
                    select(priceRange, everything()) # put priceRange as 1st column

#table(austin_validation$priceRange)

austin_test <- bake(rec_austin, new_data = test) %>% 
              select(priceRange, everything()) # put priceRange as 1st column

#table(austin_test$priceRange)

id_holdout <- holdout %>% select(uid) 
austin_holdout <- bake(rec_austin, new_data = holdout)

#table(austin_training$high_price_words)
```

Now let's create a tunable xgboost model specification, tuning a lot of the important
model hyperparameters, and combine it with our feature engineering recipe in a
`workflow()`. We can also create a custom `xgb_grid` to specify what parameters I
want to try out, like not-too-small learning rate, avoiding tree stubs, etc. I chose
this parameter grid to get reasonable performance in a reasonable amount of tuning
time.

```{r}

# dir.create("../2021-08-01-austin-r-xgb-house-price/data") # local

write_csv(austin_training, 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_training.csv", 
          col_names = FALSE)
write_csv(austin_validation, 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_validation.csv", 
          col_names = FALSE)
write_csv(austin_test %>% select(-priceRange), 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_test.csv", 
          col_names = FALSE)
write_csv(austin_test, 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_test_2.csv", 
          col_names = TRUE) # Need this later to use with results

write_csv(austin_holdout, 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_holdout.csv", 
          col_names = FALSE) 
write_csv(id_holdout, 
          "../2021-08-01-austin-r-xgb-house-price/data/austin_holdout_id.csv",
          col_names = FALSE) # Need this later to use with holdout results
```

```{r chunk13}
use_condaenv("sagemaker-r", required = TRUE)
sagemaker <- import("sagemaker")
session <- sagemaker$Session()
```

```{r}
bucket <- "twarczak-sagemaker7"
project <- "austin"    # keep short. 32 character max job_name
data_path <- paste0("s3://", bucket, "/", project, "/", "data")
models_path <- paste0("s3://", bucket, "/", project, "/", "models")
```

```{r chunk19, cache=TRUE}
s3_uploader <- sagemaker$s3$S3Uploader()

s3_train <- s3_uploader$upload(local_path = "../2021-08-01-austin-r-xgb-house-price/data/austin_training.csv", 
                               desired_s3_uri = data_path)

s3_validation <- s3_uploader$upload(local_path = "../2021-08-01-austin-r-xgb-house-price/data/austin_validation.csv",
                                    desired_s3_uri = data_path)

s3_test <- s3_uploader$upload(local_path = "../2021-08-01-austin-r-xgb-house-price/data/austin_test.csv", 
                              desired_s3_uri = data_path)
```

```{r chunk20}
region <- session$boto_region_name
container <- sagemaker$image_uris$retrieve(framework = "xgboost", 
                                           region    = region, 
                                           version   = "1.3-1" )

role_arn <- Sys.getenv("SAGEMAKER_ROLE_ARN")

xgb_estimator <- sagemaker$estimator$Estimator(image_uri          = container,
                                               role               = role_arn,
                                               instance_count     = 1L,
                                               instance_type      = "ml.m4.16xlarge", 
                                               volume_size        = 10L, 
                                               output_path        = models_path,
                                               sagemaker_session  = session,
                                               use_spot_instances = TRUE,
                                               max_run            = 1800L,
                                               max_wait           = 3600L )
                                              # keep at ml.m4.16xlarge
```

```{r chunk21}
xgb_estimator$set_hyperparameters(objective        = "multi:softprob",
                                  eval_metric      = "mlogloss",
                                  num_class        = 5L, # necessary for multiclass
                                  max_depth        = 5L, 
                                  eta              = 0.1,
                                  num_round        = 70L,
                                  colsample_bytree = 0.4,
                                  alpha            = 1L,
                                  min_child_weight = 1.1,
                                  subsample        = 0.7,
                                  gamma            = 0.01)
```

```{r chunk22}
# Create training job name based project organization principles
algo <- "xgb"    # keep short. 32 character max job_name
timestamp <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S") # timestamp
job_name_1 <- paste(project, algo, timestamp, sep = "-")
s3_train_input <- sagemaker$inputs$TrainingInput(s3_data = s3_train,
                                                 content_type = 'csv')
s3_valid_input <- sagemaker$inputs$TrainingInput(s3_data = s3_validation,
                                                 content_type = 'csv')
input_data <- list('train'      = s3_train_input,
                   'validation' = s3_valid_input)
```

```{r chunk23}
xgb_estimator$fit(inputs   = input_data,
                  job_name = job_name_1,
                  wait     = FALSE)
```

```{r chunk25}
training_job_stats <- session$describe_training_job(job_name = job_name_1)
final_metrics_1 <-  map_df(training_job_stats$FinalMetricDataList, 
                           ~tibble(metric_name = .x[["MetricName"]],
                                   value = .x[["Value"]]))
```

```{r chunk25.2}
final_metrics_1
```

```{r chunk26}
predictions_path_1 <- paste0(models_path, "/", job_name_1, "/predictions") 

xgb_batch_predictor_1 <- xgb_estimator$transformer(instance_count     = 1L, 
                                                   instance_type      = "ml.m4.16xlarge", 
                                                   strategy           = "MultiRecord",
                                                   assemble_with      = "Line",
                                                   output_path        = predictions_path_1)
```

```{r chunk27}
xgb_batch_predictor_1$transform(data         = s3_test, 
                                content_type = 'text/csv',
                                split_type   = "Line",
                                job_name     = job_name_1,
                                wait         = FALSE)
```

```{r chunk29, message=FALSE, warning=FALSE}

s3_downloader <- sagemaker$s3$S3Downloader()
s3_test_predictions_path_1 <- s3_downloader$list(predictions_path_1)
 
dir.create("./predictions")
s3_downloader$download(s3_test_predictions_path_1, "./predictions")
 

test_predictions_1 <- read_csv("./predictions/austin_test.csv.out",
                                col_names = FALSE  ) #%>% 
                      # pull(X1)
tp1 <- unnest(test_predictions_1)

test_predictions_2 <- tp1 %>% 
                      transmute("0-250000"      = X1,
                                "250000-350000" = X2,
                                "350000-450000" = X3,
                                "450000-650000" = X4,
                                "650000+"       = X5 )

test_predictions_2$`0-250000` <- as.numeric(str_replace_all(test_predictions_2$`0-250000`,"\\[|\\]", ""))
test_predictions_2$`650000+` <- as.numeric(str_replace_all(test_predictions_2$`650000+`,"\\[|\\]", "")) 

test_results_1 <- austin_test %>% 
  select(priceRange) %>% 
  bind_cols(test_predictions_2) %>% 
  rename(Truth = priceRange) %>% 
  mutate(Truth = case_when(Truth == "0" ~ "0-250000", 
                           Truth == "1" ~ "250000-350000",
                           Truth == "2" ~ "350000-450000",
                           Truth == "3" ~ "450000-650000", 
                           Truth == "4" ~ "650000+"      ))
```

```{r chunk29.2}
head(test_results_1)
```

```{r message=FALSE, warning=FALSE, fig.height=6, fig.width=8}
test_results_1 <- test_results_1 %>%
    mutate(pred_class = pmap_chr(across(`0-250000`:`650000+`),
                        ~ names(c(...)[which.max(c(...))])))

conf1 <- test_results_1 %>% 
  conf_mat(Truth, pred_class) %>% 
  autoplot(type = "heatmap") +
  scale_fill_scico(begin = 0.2, end = 0.7, palette = "bamako") +  
  labs(title = "Confusion Matrix - XGBoost multi-class BEFORE Tuning\n") +
  scale_x_discrete(position = "top")
conf1
```


```{r chunk30, message=FALSE, warning=FALSE, cache=TRUE, fig.width=10}
roc1 <- test_results_1 %>% 
  roc_curve(as.factor(Truth), `0-250000`:`650000+`) %>% 
  ggplot(aes(1-specificity, sensitivity, color = .level)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(alpha = 0.7, size = 1.1) +
  coord_equal() +
  labs(color = "priceRange",
       title = "ROC curves for XGBoost multi-class",
       subtitle = "AUC = 0.8684319") +
  theme(legend.position = c(0.8, 0.3),
        legend.background = element_rect(fill = "black"))
roc1
```

```{r}
# res <- test_results_1 %>% roc_curve(as.factor(Truth), `0-250000`:`650000+`)
# res
```

```{r chunk32}
xgb_estimator <- sagemaker$estimator$Estimator(image_uri          = container,
                                               role               = role_arn,
                                               instance_count     = 1L,
                                               instance_type      = "ml.m5.4xlarge", 
                                               volume_size        = 10L, 
                                               output_path        = models_path,
                                               sagemaker_session  = session,
                                               use_spot_instances = TRUE,
                                               max_run            = 1800L,
                                               max_wait           = 3600L )
                                              # keep at ml.m4.16xlarge

xgb_estimator$set_hyperparameters(objective        = "multi:softprob",
                                  eval_metric      = "mlogloss",
                                  num_class        = 5L,
                                  alpha            = 1L ) 
#max_depth 0 w/ tree_method "hist" & grow_policy "lossguide"
                                 # tree_method      = "hist",
                                 # grow_policy      = "lossguide",
```
**Tunable**
```{r chunk33}
hyperp_ranges <- list(num_round = sagemaker$tuner$IntegerParameter(500L, 800L),
                      max_depth = sagemaker$tuner$IntegerParameter(5L, 8L),
              eta              = sagemaker$tuner$ContinuousParameter(0.01,0.07),
              colsample_bytree = sagemaker$tuner$ContinuousParameter(0.5, 0.75),
              min_child_weight = sagemaker$tuner$ContinuousParameter(1.0, 7.0),
              subsample        = sagemaker$tuner$ContinuousParameter(0.5, 0.9),
              gamma            = sagemaker$tuner$ContinuousParameter(0.01, 0.1))
```
Create `HyperparameterTuner` object.
```{r chunk34}
tuner <- sagemaker$tuner$HyperparameterTuner(estimator             = xgb_estimator,
                                             objective_metric_name = "validation:mlogloss",
                                             objective_type        = "Minimize",
                                             hyperparameter_ranges = hyperp_ranges, 
                                             strategy              = "Bayesian",
                                             max_jobs              = 150L,
                                             max_parallel_jobs     = 4L)
```
Define S3 location of data sets and tuning job name.
```{r chunk35}
algo <- "xgb"
timestamp <- format(Sys.time(), "%Y-%m-%d-%H-%M-%S")
job_name_x <- paste(project, algo, timestamp, sep = "-")
s3_train_input <- sagemaker$inputs$TrainingInput(s3_data = s3_train,
                                                 content_type = 'csv')
s3_valid_input <- sagemaker$inputs$TrainingInput(s3_data = s3_validation,
                                                 content_type = 'csv')
input_data <- list('train'      = s3_train_input,
                   'validation' = s3_valid_input)
```
### Step 9 - Start tuning job
```{r chunk36}
tuner$fit(inputs   = input_data, 
          job_name = job_name_x,
          wait     = FALSE )
```
### Step 10 - Evaluate tuning job results
```{r chunk38}
# find the jobs names in aws services: sagemaker, training, hyperparameter tuning jobs
tuning_job_results <- sagemaker$HyperparameterTuningJobAnalytics("austin-xgb-2021-08-21-19-26-19")
tuning_results_df <- tuning_job_results$dataframe()
```

```{r chunk38.2}
tuning_results_df %>% arrange(FinalObjectiveValue) %>% select(FinalObjectiveValue, everything())
```
Plot a time series chart that shows how AUC developed over 100 training jobs, tuned
by the Bayesian optimizer.
```{r message=FALSE, warning=FALSE}
tune_plot0 <-ggplot(tuning_results_df, aes(TrainingEndTime, FinalObjectiveValue)) +
  geom_point() +
  xlab("Time") +
  ylab(tuning_job_results$description()$TrainingJobDefinition$StaticHyperParameters$`_tuning_objective_metric`) +
  ggtitle("Hyperparameter tuning mLogloss",  
          "Progression over the period of all 100 training jobs") 

tune_plot1 <- ggplot(tuning_results_df, aes(num_round, max_depth)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_scico("mLogloss", palette = "roma", direction = 1) +
  ggtitle("Hyperparameter tuning mLogloss", 
          "Using a Bayesian strategy") 

tune_plot2 <-ggplot(tuning_results_df, aes(eta, colsample_bytree)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_scico("mLogloss", palette = "roma", direction = 1) +
  ggtitle("Hyperparameter tuning mLogloss", 
          "Using a Bayesian strategy") 

tune_plot3 <-ggplot(tuning_results_df, aes(min_child_weight, subsample)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_scico("mLogloss", palette = "roma", direction = 1) +
  ggtitle("Hyperparameter tuning mLogloss", 
          "Using a Bayesian strategy") 

tune_plot4 <-ggplot(tuning_results_df, aes(gamma, eta)) +
  geom_point(aes(color = FinalObjectiveValue)) +
  scale_color_scico("mLogloss", palette = "roma", direction = 1) +
  ggtitle("Hyperparameter tuning mLogloss", 
          "Using a Bayesian strategy") 
```

```{r message=FALSE, warning=FALSE, fig.height=10, fig.width=12}
(tune_plot0 + tune_plot1) / (tune_plot2 + tune_plot3) / (tune_plot4 + plot_spacer())
```

```{r chunk41, message=FALSE, warning=FALSE}
best_5_tuned_models <- tuning_results_df %>%
                    #filter(FinalObjectiveValue == max(FinalObjectiveValue)) %>%
                    arrange(FinalObjectiveValue) %>% 
                    head(n=5) 

#unnest(best_10_tuned_models) %>% select(FinalObjectiveValue, everything())

best_tuned_model <- tuning_results_df %>%
                    #filter(FinalObjectiveValue == max(FinalObjectiveValue)) %>%
                    arrange(FinalObjectiveValue) %>% 
                    head(n=1) %>% pull(TrainingJobName)

#best_tuned_model
training_job_stats <- session$describe_training_job(job_name = best_tuned_model)

final_metrics_2 <-  map_df(training_job_stats$FinalMetricDataList, 
                           ~tibble(metric_name = .x[["MetricName"]],
                                   value       = .x[["Value"]]))
```

```{r chunk41.2, cache=TRUE}
final_metrics_2
```

```{r chunk42, eval=FALSE}
predictions_path_2 <- paste0(models_path, "/", best_tuned_model, "/predictions")

session$create_model_from_job(best_tuned_model)

xgb_batch_predictor_2 <- sagemaker$transformer$Transformer(model_name     = best_tuned_model,
                                                           instance_count = 1L, 
                                                           instance_type  = "ml.m4.4xlarge", 
                                                           strategy       = "MultiRecord",  
                                                           assemble_with  = "Line",
                                                           output_path    = predictions_path_2)
```
### Step 12 - Start batch prediction job

```{r chunk43}
xgb_batch_predictor_2$transform(data         = s3_test, 
                                content_type = 'text/csv',
                                split_type   = "Line",
                                job_name     = "austin-xgb-2021-08-20-22-08-07-083-347756be",
                                wait         = FALSE) 
```


```{r chunk45}
s3_downloader <- sagemaker$s3$S3Downloader()

s3_test_predictions_path_2 <- s3_downloader$list(predictions_path_2)
 
#dir.create("./predictions")

s3_downloader$download(s3_test_predictions_path_2, "./predictions")
 
test_predictions_2 <- read_csv("./predictions/austin_test.csv.out",
                               col_names = FALSE) #%>% pull(X1)

tp2 <- unnest(test_predictions_2)

test_predictions_3 <- tp2 %>%
                      transmute("0-250000"      = X1,
                                "250000-350000" = X2,
                                "350000-450000" = X3,
                                "450000-650000" = X4,
                                "650000+"       = X5 )
 
test_predictions_3$`0-250000` <- as.numeric(str_replace_all(test_predictions_3$`0-250000`,"\\[|\\]", ""))
test_predictions_3$`650000+` <- as.numeric(str_replace_all(test_predictions_3$`650000+`,"\\[|\\]", "")) 

austin_test <- read_csv("./data/austin_test_2.csv")

test_results_2 <- austin_test %>%
  select(priceRange) %>%
  bind_cols(test_predictions_3) %>%
  rename(Truth = priceRange) %>%
  mutate(Truth = case_when(Truth == "0" ~ "0-250000",
                           Truth == "1" ~ "250000-350000",
                           Truth == "2" ~ "350000-450000",
                           Truth == "3" ~ "450000-650000",
                           Truth == "4" ~ "650000+"      ))
```

```{r chunk45.2, cache=TRUE}

```

```{r message=FALSE, warning=FALSE}
test_results_2 <- test_results_2 %>%
    mutate(pred_class = pmap_chr(across(`0-250000`:`650000+`),
                        ~ names(c(...)[which.max(c(...))])))
```

```{r message=FALSE, warning=FALSE, fig.height=6, fig.width=14}
conf2 <- test_results_2 %>% 
  conf_mat(Truth, pred_class) %>% 
  autoplot(type = "heatmap") +
  scale_fill_scico(begin = 0.2, end = 0.7, palette = "bamako") +  
  labs(title = "Confusion Matrix - XGBoost multi-class AFTER Tuning\n") +
  scale_x_discrete(position = "top")
# compare to initial model

conf1 + conf2
```
```{r}
conf1
```

```{r message=FALSE, warning=FALSE, include=FALSE}
# if we wanted to use multi_roc

# test_results_2_auc <- test_results_2 %>% 
# mutate(Truth = factor(case_when(Truth == "0-250000"      ~ "P1_true",
#                                 Truth == "250000-350000" ~ "P2_true",
#                                 Truth == "350000-450000" ~ "P3_true",
#                                 Truth == "450000-650000" ~ "P4_true",
#                                 Truth == "650000+"       ~ "P5_true" ),
#                       levels = c("P1_true", "P2_true", "P3_true", "P4_true", "P5_true")),
#        n = 1) %>%
# rename(P1_pred_m1 = "0-250000",
#        P2_pred_m1 = "250000-350000",
#        P3_pred_m1 = "350000-450000",
#        P4_pred_m1 = "450000-650000",
#        P5_pred_m1 = "650000+" ) %>% 
# mutate(row_num = 1:n()) %>% 
# pivot_wider(names_from = Truth, values_from = n, values_fill = 0) %>% 
# select(-c(row_num, pred_class))
# 
# auc_mult <- multi_roc(data.frame(test_results_2_auc), force_diag = TRUE) # need df, not tibble
# auc_mult$AUC
# plot_roc_df <- plot_roc_data(auc_mult)
# cal_auc()
# multiclass.roc(test_results_2$Truth, test_results_2$pred_class)
# 
# roc3 <- plot_roc_df %>% 
#   ggplot(aes(1-Specificity, Sensitivity)) +
#   geom_abline(lty = 2, color = "gray80", size = 1.5) +
#   geom_path(aes(color = Group), alpha = 0.7, size = 1.1) +
#   coord_equal() +
#   labs(color = "priceRange",
#        title = "ROC curves for XGBoost multi-class after Tuning") 
```

```{r chunk30, message=FALSE, warning=FALSE, cache=TRUE, fig.height=6, fig.width=12}
auc2 <- test_results_2 %>% 
  roc_auc(as.factor(Truth), `0-250000`:`650000+`)
auc1 <- test_results_1 %>% 
  roc_auc(as.factor(Truth), `0-250000`:`650000+`)

roc2 <- test_results_2 %>% 
  roc_curve(as.factor(Truth), `0-250000`:`650000+`) %>% 
  ggplot(aes(1-specificity, sensitivity, color = .level)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(alpha = 0.7, size = 1.1) +
  coord_equal() +
  labs(color = "priceRange",
       title = "ROC curves for XGBoost multi-class AFTER Tuning",
       subtitle = "AUC = 0.8821398") +
  theme(legend.position = c(0.8, 0.3),
        legend.background = element_rect(fill = "black"))

roc1 + roc2
```

```{r }
#library(SHAPforxgboost)
#library(xgboost)
#library(data.table)

unnest(best_5_tuned_models) %>% select(c(FinalObjectiveValue, 1:7)) # final obj = mlogloss

param_list <- list(objective = "multi:softprob",
                   eval_metric = "mlogloss",
                   num_class = 5,
                   eta = 0.03411075,
                   max_depth = 6,
                   gamma = 0.01,
                   subsample = 0.7702243,
                   colsample_bytree = 0.6781905,
                   min_child_weight = 3.299003 )

austin_training_2 <- data.table(austin_training)
dataX_2 <- as.matrix(austin_training_2[,-1])

mod <- xgboost::xgboost(data = dataX_2, 
                        label = as.matrix(austin_training_2[["priceRange"]]), 
                        params = param_list, nrounds = 509,
                        verbose = FALSE, nthread = parallel::detectCores() - 2,
                        early_stopping_rounds = NULL)

head(dataX_2)
```

```{r}
shap_values <- shap.values(xgb_model = mod, X_train = dataX_2) # takes a while

shap_values$mean_shap_score

shap_long <- shap.prep(xgb_model = mod, X_train = dataX_2) # takes a while
```


```{r fig.width=10, fig.height=6}
shap.plot.summary(shap_long) +
  ggdark::dark_theme_bw() +
  theme(legend.position = 'bottom') +
  labs(title = "SHAP Values for Features in XGBoost Model")
```
```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
shap_force_plot <- shap.prep.stack.data(shap_contrib = shap_values$shap_score, 
                                        top_n = 3, n_groups = 6)
# you may choose to zoom in at a location, and set y-axis limit using `y_parent_limit`  
shap.plot.force_plot(shap_force_plot, zoom_in_group = 5, y_parent_limit = c(-2.2,2.2)) +
  ggdark::dark_theme_bw() +
  labs(title = "SHAP Force Plot",
       subtitle = "Shows feature contribution to XGBoost prediction at each observation")

```
```{r fig.height=10, fig.width=10}
shap.plot.force_plot_bygroup(shap_force_plot) +
  ggdark::dark_theme_bw() +
  labs(title = "SHAP Force Plot Clusters - Austin Neighborhoods?")
```
```{r}
best_tuned_model
```

```{r chunk49, eval=FALSE}
boto_client <- session$boto_session$client("sagemaker")
austin_model <- boto_client$list_models()[["Models"]] %>% 
  map_chr("ModelName") %>% 
  .[[1]]
```

```{r chunk49.2}
austin_model
```

```{r}
```


## Batch Inference on Holdout Data

We needed to submit probabilities of every `priceRange` bin for each observation (home). There are 5 possible price bins:
`0-250000`, `250000-350000`, `350000-450000`, `450000-650000`, `650000+`
So we'll need to submit a .cvs file containing the holdout id column + 5 price columns containing our predicted probabilities. To run inference we'll pass our "baked" holdout data into our best performing XGboost model.  
```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```


```{r}
```






```{r}
xgb_spec <- 
  boost_tree(trees = 1000, 
             tree_depth = tune(),
             min_n = tune(),
             mtry = tune(),
             sample_size = tune(),
             learn_rate = tune()) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

xgb_word_wf <- workflow(austin_rec, xgb_spec)

set.seed(123)
xgb_grid <-
  grid_max_entropy(
    tree_depth(c(5L, 10L)),
    min_n(c(10L, 40L)),
    mtry(c(5L, 10L)),
    sample_prop(c(0.5, 1.0)),
    learn_rate(c(-2, -1)),
    size = 20
  )

xgb_grid
```

Now we can tune across the grid of parameters and our resamples. Since we are trying
quite a lot of hyperparameter combinations, let's use
[racing](https://juliasilge.com/blog/baseball-racing/) to quit early on clearly bad
hyperparameter combinations.

```{r}
library(finetune)
doParallel::registerDoParallel()

set.seed(234)
xgb_word_rs <- 
  tune_race_anova(
    xgb_word_wf,
    austin_folds,
    grid = xgb_grid,
    metrics = metric_set(mn_log_loss),
    control = control_race(verbose_elim = TRUE)
  )

xgb_word_rs
```

That takes a little while but we did it!

## Evaluate results

First off, how did the "race" go?

```{r}
plot_race(xgb_word_rs)
```

We can look at the top results manually as well.

```{r}
show_best(xgb_word_rs)
```

Let's use `last_fit()` to fit one final time to the **training** data and evaluate
one final time on the **testing** data, with the numerically optimal result from
`xgb_word_rs`.

```{r}
xgb_last <- 
  xgb_word_wf %>%
  finalize_workflow(select_best(xgb_word_rs, "mn_log_loss")) %>%
  last_fit(austin_split)

xgb_last
```

How did this model perform on the testing data, that was not used in tuning/training?

```{r}
collect_predictions(xgb_last) %>%
    mn_log_loss(priceRange, `.pred_0-250000`:`.pred_650000+`)
```

This result is pretty good for a single (not ensembled) model and is a wee bit better
than what I did during the SLICED competition. I had an R bomb right as I was
finishing up tuning a model just like the one I am demonstrating here!

How does this model perform across the different classes?

```{r}
collect_predictions(xgb_last) %>%
    conf_mat(priceRange, .pred_class) %>%
    autoplot()
```

We can also visualize this with an ROC curve.

```{r}
collect_predictions(xgb_last) %>%
    roc_curve(priceRange,  `.pred_0-250000`:`.pred_650000+`) %>%
    ggplot(aes(1 - specificity, sensitivity, color = .level)) +
    geom_abline(lty = 2, color = "gray80", size = 1.5) +
    geom_path(alpha = 0.8, size = 1.2) +
    coord_equal() +
    labs(color = NULL)
```

Notice that it is easier to identify the most expensive homes but more difficult to
correctly classify the less expensive homes.

What features are most important for this xgboost model?

```{r}
library(vip)
extract_workflow(xgb_last) %>%
    extract_fit_parsnip() %>%
    vip(geom = "point", num_features = 15)
```

The spatial information in latitude/longitude are by far the most important. Notice
that the model uses `low_price_words` more than it uses, for example, whether there
is a spa or whether it is a single family home (as opposed to a townhome or condo).
It looks like the model is trying to distinguish some of those lower priced
categories. The model does *not* really use the `high_price_words` variable, perhaps
because it is already easy to find the expensive houses.

The two finalists from SLICED go on to compete next Tuesday, which should be fun and
interesting to watch! I have enjoyed the opportunity to participate this season.
